---
title: "VAERS_22022"
author: "Prasad Nishtala"
date: "22/02/2022"
output:
  html_document: default
  pdf_document: default
---

## R Markdown

This the document to analyse data from Jan 2021 to Dec 2021 downloaded from https://vaers.hhs.gov/data/datasets.html?.

#1 **load packages and setwd**

```{r echo=TRUE}
setwd("C:/Users/pn403/OneDrive - University of Bath/Desktop/Prasad$Nishtala/vaers_2222/data")
library(data.table)
library(rio)
library(anytime)
library(readr)
library(readxl)
library(ggtext)
library(glue)
library(broom)
library(tidyverse)
library(tidyr)
library(dplyr)
library(purrr)
options(warn = -1)
```

#2 ***create a data file for vax and change coulmn VAERS_ID to numeric**

```{r echo=TRUE, warning=TRUE}
vaers_vax1 <- import("data/2021VAERSVAX.csv")

vaers_vax1$VAERS_ID <- as.numeric(vaers_vax1$VAERS_ID)

vaers_vax2 <- import("data/2022VAERSVAX.csv")

vaers_vax2$VAERS_ID <- as.numeric(vaers_vax2$VAERS_ID)

vaers_vax <- rbind(vaers_vax1,vaers_vax2)

save(vaers_vax, file = "rdata/vaers_vax.RData")

```

#3 ***create a data file for symptoms**

```{r include=FALSE}
vaers_symptoms1 <- import("data/2021VAERSSYMPTOMS.csv")

vaers_symptoms1$VAERS_ID <- as.numeric(vaers_symptoms1$VAERS_ID)

vaers_symptoms2 <- import("data/2022VAERSSYMPTOMS.csv")

vaers_symptoms2$VAERS_ID <- as.numeric(vaers_symptoms2$VAERS_ID)

vaers_symptoms <- rbind(vaers_symptoms1,vaers_symptoms2)

save(vaers_symptoms, file = "rdata/vaers_symptoms.RData")
```

#4 **create a data file for data**
```{r include=FALSE}
vaers_data1 <- import("data/2021VAERSData.csv")

vaers_data1$VAERS_ID <- as.numeric(vaers_data1$VAERS_ID)

vaers_data2 <- import("data/2022VAERSData.csv")

vaers_data2$VAERS_ID <- as.numeric(vaers_data2$VAERS_ID)

vaers_data <- rbind(vaers_data1,vaers_data2)

vaers_data %>% mutate_at(
  vars(
    "RECVDATE",
    "RPT_DATE",
    "DATEDIED",
    "VAX_DATE",
    "ONSET_DATE",
    "TODAYS_DATE"
  ),
  as.Date,
  format = "%Y-%m-%d"
)
  
save(vaers_data, file = "rdata/vaers_data.RData")
```

#5 ***load all files of vaers**

```{r message=TRUE, warning=TRUE, include=FALSE, paged.print=TRUE}
setwd("C:/Users/pn403/OneDrive - University of Bath/Desktop/vaers2222/rdata")
load("vaers_data.Rdata")
load("vaers_symptoms.Rdata")
load("vaers_vax.Rdata")

vaers_data <- vaers_data %>%
  rename_all(tolower) %>%
  select(
    -c(
      cage_yr,
      cage_mo,
      datedied,
      l_threat,
      rpt_date,
      hospital,
      hospdays,
      x_stay,
      lab_data,
      v_adminby,
      splttype,
      form_vers,
      ofc_visit,
      birth_defect,
      died,
      disable,
      er_visit
    )
  )

vaers_symptom <- vaers_symptoms %>%
  rename_all(tolower) %>%
  select(
    -c(
      symptomversion1,
      symptomversion2,
      symptomversion3,
      symptomversion4,
      symptomversion5
    )
  )

vaers_vax <- vaers_vax %>%
  rename_all(tolower) %>%
  select(-vax_lot)

```
#6 **inner_join and get a dataset with duplicates**

```{r include=FALSE}
vaers_final<-inner_join(vaers_data, vaers_symptom,by="vaers_id") %>% 
  inner_join(.,vaers_vax,by="vaers_id") %>% 
  select(-v_fundby)
```
#7 remove duplicates -dedepulicated dataset

```{r}
vaers_final_distinct <-
  inner_join(vaers_data, vaers_symptom, by = "vaers_id") %>%
  inner_join(., vaers_vax, by = "vaers_id") %>%
  select(-v_fundby) %>%
  filter(vax_type == "COVID19") %>%
  distinct(vaers_id, vax_date, state, age_yrs, sex, .keep_all = TRUE)

```
##8 **create table for study population**
```{r}
library(gtsummary)
reset_gtsummary_theme()
theme_gtsummary_compact()
vaers_final_distinct %>%
  select(c("age_yrs", "sex", "recovd", "numdays", "vax_manu")) %>%
  rename(
    "Age" = age_yrs,
    "Sex" = sex,
    "Recovered" = recovd,
    "Days to symptom onset" = numdays,
    "Vaccine manufacturer" = vax_manu
  ) %>%
  tbl_summary(by = "Vaccine manufacturer") %>%
  add_stat_label() %>%
  add_overall() %>%
  bold_labels() %>%
  as_gt() %>%
  gt::tab_header("Characteristics of the study population")
```

#9   **get total adrs by vaccine type**
```{r}
vaers_total_adrs <- vaers_final_distinct %>%
  dplyr::filter(vax_manu=="MODERNA") %>% 
  select(vaers_id,symptom1:symptom5) %>%
  pivot_longer(-vaers_id,names_to = "symptoms", values_to = "adr") %>%
  distinct(vaers_id, symptoms, adr, .keep_all = TRUE) %>%
  count(adr, sort = TRUE)
```
#10 plot adrs

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
#read_chunk('C:/Users/pn403/OneDrive - University of #Bath/Desktop/minimalR/vaers_2222/plot_adrs_moderna_5000.R')
library(ggplot2)
library(shadowtext)
library(grid)

vaers_total_adrs$adr <- as.factor(vaers_total_adrs$adr)
plt <- vaers_total_adrs %>%
  filter(!adr=="") %>%
  filter(n>5000) %>%
ggplot+
  geom_col(aes(x=n, y=adr), width = 0.9)

plt <- plt +
  scale_x_continuous(
    limits = c(0, 50000),
    breaks = seq(5000, 50000, by = 10000),
    expand = c(0, 0), # The horizontal axis does not extend to either side
    position = "top"  # Labels are located on the top
  ) +
  # The vertical axis only extends upwards
  scale_y_discrete(expand = expansion(add = c(0, 0.5))) +
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "sans", size = 12)
  )


plt2 <- plt +
  geom_shadowtext(
    #data = subset(data, n < 5000),
    aes(n, y = adr, label = adr),
    hjust = 0,
    nudge_x = 0.3,
    bg.colour = "white",
    bg.r = 0.05,
    family = "sans",
    size = 2.5
  ) +
  geom_text(
    #data = subset(data, n >= 5000),
    aes(0, y = adr, label = adr),
    hjust = 0,
    nudge_x = 0.3,
    colour = "white",
    family = "sans",
    size = 2.5
  )


plt3 <- plt2 +
  labs(
    title = "Moderna",
    subtitle = "Adverse drug events > 5000 from case reports, 2021-2022"
  ) +
  theme(
    plot.title = element_text(
      family = "sans",
      face = "bold",
      size = 22
    ),
    plot.subtitle = element_text(
      family = "sans",
      size = 20
    )
  )

# Make room for annotations
plt4 <- plt3 +
  theme(
    plot.margin = margin(0, 0, 0, 0.0, "npc")
  )

# Print the ggplot2 plot

# Add horizontal line on top
# It goes from x = 0 (left) to x = 1 (right) on the very top of the chart (y = 1)
# You can think of 'gp' and 'gpar' as 'graphical parameters'.
# There we indicate the line color and width
grid.lines(
  x = c(0, 1),
  y = 1,
  gp = gpar(col = "#e5001c", lwd = 4)
)

# Add rectangle on top-left
# lwd = 0 means the rectangle does not have an outer line
# 'just' gives the horizontal and vertical justification
grid.rect(
  x = 0,
  y = 1,
  width = 0.05,
  height = 0.025,
  just = c("left", "top"),
  gp = gpar(fill = "#e5001c", lwd = 0)
)

print(plt4)


```

#11 **chills as example to get some exploratory values**

```{r}
chills<-vaers_final_distinct %>% 
  select(vaers_id,symptom1:symptom5,vax_type) %>% 
  filter_at(vars(symptom1:symptom5), any_vars(. == 'Chills')) %>% 
  count () %>% 
  print()
```
#12
```{r}
chills_by_vaxmanu <- vaers_final_distinct %>%
  select(symptom1:symptom5, vax_manu) %>%
  filter_at(vars(symptom1:symptom5), any_vars(. == "Chills")) %>%
  group_by(vax_manu) %>%
  count () %>%
  print()
```

#13**total adrs**

```{r}
  vaers_final_distinct <-
  inner_join(vaers_data, vaers_symptom, by = "vaers_id") %>%
  inner_join(., vaers_vax, by = "vaers_id") %>%
  select(-v_fundby) %>%
  filter(vax_type == "COVID19") %>%
  distinct(vaers_id, vax_date, state, age_yrs, sex, .keep_all = TRUE) %>%
  count() %>%
  pull(n[]) %>%
  as.integer() %>%
  print()

```
#14***stats**
##*a number of adrs of interest with vaccine type*
```{r echo=TRUE, error=TRUE, message=FALSE, warning=FALSE}
a<-chills_by_vaxmanu %>% filter(vax_manu=="MODERNA")
a<-sum(a$n)
print(a)

##number of other adrs with vaccine type ( b value)
vaers_final_distinct <-
  inner_join(vaers_data, vaers_symptom, by = "vaers_id") %>%
  inner_join(., vaers_vax, by = "vaers_id") %>%
  select(-v_fundby) %>%
  filter(vax_type == "COVID19") %>%
  distinct(vaers_id, vax_date, state, age_yrs, sex, .keep_all = TRUE)

  b <- vaers_final_distinct %>% 
  select(symptom1:symptom5, vax_manu) %>%
  filter(vax_manu == "MODERNA") %>%
  filter(if_all(everything(), ~ !str_detect(., "Chills"))) %>%
  group_by(vax_manu) %>%
  count () %>%
  pull(n) %>%
  print()

##number if adrs of interest with all other vaccine type ( c value)

c<-chills_by_vaxmanu %>% filter(vax_manu!="MODERNA")
c<-sum(c$n)
print(c)


##number of all other adrs with all other vaccines types except adr of interest (d value)

d <- vaers_final_distinct %>%
  select(symptom1:symptom5, vax_manu) %>%
  filter(vax_manu != "MODERNA") %>%
  filter(if_all(everything(), ~ !str_detect(., "Chills"))) %>%
  # group_by(vax_manu) %>%
  count () %>%
  pull(n) %>%
  print()
```

##plot for inflammation >25
```{r echo=TRUE, error=TRUE, message=FALSE, warning=FALSE}
vaers_total_adrs$adr <- as.factor(vaers_total_adrs$adr)
plt <- vaers_total_adrs %>%
  filter(!adr=="") %>%
  filter(grepl("itis",adr)& n>25) %>%
ggplot+
  geom_col(aes(x=n, y=adr), width = 0.8)

plt <- plt +
  scale_x_continuous(
    limits = c(0, 1000),
    breaks = seq(0, 1000, by = 100),
    expand = c(0, 0), # The horizontal axis does not extend to either side
    position = "top"  # Labels are located on the top
  ) +
  # The vertical axis only extends upwards
  scale_y_discrete(expand = expansion(add = c(0, 0.5))) +
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # Remove labels from the vertical axis
    axis.text.y = element_blank(),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "sans", size = 12)
  )

plt

plt2 <- plt +
  geom_shadowtext(
    #data = subset(data, n < 5000),
    aes(n, y = adr, label = adr),
    hjust = 0,
    nudge_x = 0.3,
    bg.colour = "white",
    bg.r = 0.05,
    family = "sans",
    size = 2.5
  ) +
  geom_text(
    #data = subset(data, n >= 5000),
    aes(0, y = adr, label = adr),
    hjust = 0,
    nudge_x = 0.3,
    colour = "white",
    family = "sans",
    size = 2.5
  )

plt2

plt3 <- plt2 +
  labs(
    title = "Moderna",
    subtitle = "Adverse drug events related to inflammation n>25 from case reports, 2021"
  ) +
  theme(
    plot.title = element_text(
      family = "sans",
      face = "bold",
      size = 22
    ),
    plot.subtitle = element_text(
      family = "sans",
      size = 18
    )
  )
plt3


# Make room for annotations
plt4 <- plt3 +
  theme(
    plot.margin = margin(0.09, 0, 0.02, 0.02, "npc")
  )

# Print the ggplot2 plot
plt4

# Add horizontal line on top
# It goes from x = 0 (left) to x = 1 (right) on the very top of the chart (y = 1)
# You can think of 'gp' and 'gpar' as 'graphical parameters'.
# There we indicate the line color and width
grid.lines(
  x = c(0, 1),
  y = 1,
  gp = gpar(col = "#e5001c", lwd = 4)
)

# Add rectangle on top-left
# lwd = 0 means the rectangle does not have an outer line
# 'just' gives the horizontal and vertical justification
grid.rect(
  x = 0,
  y = 1,
  width = 0.05,
  height = 0.025,
  just = c("left", "top"),
  gp = gpar(fill = "#e5001c", lwd = 0)
)

plt4


```

